---
title: "Lab 5"
output:
  - github_document
  - html_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(tidyverse)
```


```{r}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```
```{r}

if(!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", 
              timeout = 60)
}
met <- data.table::fread("met_all.gz")
```


```{r}
#Merge the data 

met <- merge( 
  x = met, 
  y = stations,
  all.x = TRUE, all.y = FALSE,
  by.x = "USAFID", by.y = "USAF"
  )
```

Question 1: Representative station for the US
```{r}
station_averages <- met[,. (
  temp = mean(temp, na.rm = TRUE), 
  wind.sp = mean(wind.sp, na.rm= TRUE),
  atm.press = mean(atm.press, na.rm= TRUE),
  lat = mean(lat, na.rm= TRUE),
  lon = mean(lon, na.rm= TRUE)
), by= USAFID]
```

```{r}
medians <- station_averages[,.(
  temp_50= quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50= quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm.press_50= quantile(atm.press, probs = .5, na.rm = TRUE)
)]

medians
```
Now we can find stations that are closest to these 
```{r}
station_averages[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station
```
The median temperature station is `r median_temp_station $ USAFID`

Question 2: Representative station per state

We need to first recover the state variable by merging
```{r}
station_averages <- merge(x = station_averages, y = stations, by.x = "USAFID", by.y = "USAF", 
    all.x = TRUE, all.y = FALSE)
```

Now we can compute the median per state 
```{r}

station_averages[, temp_50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, wind.sp_50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE]
```
Now the euclidean distance... 
```{r}
station_averages[, eudist := sqrt(
  (temp - temp_50)^2 + (wind.sp - wind.sp_50)^2
  )]
station_averages
```
Select unique median station per state at the lowest latitude...

```{r} 

#Identify station with the lowest latitude in each state
median_station_state <- station_averages[, list(USAFID, lat, lon, STATE)]
lat_station <-  median_station_state[ , list(USAFID, lat_min = min(lat), lat, lon), by = STATE]
lat_station[lat_min == lat]

lat_station

```


Question 3: In the middle? Visualize station closest to the midpoint of the state + representative station using leaflet 

Compute midpoint per state 
```{r}

station_averages[, lat_mdpt := (max(lat) + min(lat)) / 2, by = STATE]
station_averages[, lon_mdpt := (max(lon) + min(lon)) / 2, by = STATE]

station_averages[, lat_distance := abs(lat - lat_mdpt)]
station_averages[, lon_distance := abs(lon - lon_mdpt)]

station_mdpt <-  station_averages[, list(USAFID, lat, lon, lat_distance, lon_distance, hyp = sqrt(lat_distance^2 + lon_distance^2)), by = STATE]
station_hyp <- station_mdpt[, list(USAFID, lat, lon, hyp_min = min(hyp), hyp), by = STATE]
station_hyp[hyp_min == hyp]

station_hyp
```

Leaflet Plot
```{r}


library(leaflet)
#leaflet(lat_station) %>%
#  addProviderTiles('CartoDB.Positron') %>%
#  addCircles(lat = ~lat, lng = ~lon, opacity = 1, fillOpacity = 1, radius = 400)

leaflet() %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
  data = lat_station[lat_min == lat],
  lat=~lat, lng=~lon,
  opacity = 1, fillOpacity = 1, radius = 10, color = "blue") %>%
  addCircles(
  data = station_hyp[hyp_min == hyp],
  lat=~lat, lng=~lon,
  opacity = 1, fillOpacity = 1, radius = 10, color = "red")
```


Question 4: Means of means

```{r}
met[, state_temp := mean(temp, na.rm = TRUE), by = STATE]
met[, temp_cat := fifelse(
  state_temp < 20, "low-temp",
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]
```

Let's make sure we don't have NAs

```{r}
table(met$temp_cat, useNA = "always")
```

Include all other variables....
```{r}
tab <- met[, .(
  N_entries = .N,
  N_stations = length(unique(USAFID)),
  N_states = length(unique(STATE)),
  mean_temp = mean(temp, na.rm = TRUE),
  mean_wind.sp = mean(wind.sp, na.rm= TRUE),
  mean_atm.pres = mean(atm.press, na.rm= TRUE)
) ,by = temp_cat]

knitr::kable(tab) 
```

