---
title: "Lab 6 10-1-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

```{r}
library(data.table)
library(ggplot2)
library(tidyverse)
library(tidytext)
```

Download the data
```{r}
fn <- "mtsamples.csv"
if(!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", destfile = fn)

mtsamples <- read.csv(fn) #reads data into a data frame from base R 
mtsamples <- as_tibble(mtsamples)
```

1- What specialties do we have? 
```{r}
library(dplyr)
specialities <- mtsamples %>% count(medical_specialty) 

#Number of different medical specialties
specialities
nrow(specialities)

#histogram
specialities %>%
ggplot(aes(n, fct_reorder(medical_specialty, n))) + 
  geom_col() 

#Table of top 15 specialties 
specialities %>% 
  arrange(desc(n)) %>%
  top_n(n, 15) %>%
  knitr::kable()
  
```
There are `r nrow(specialities)` of different medical specialties. They are not evenly distributed, as the majority are surgeons and then less than half are consultants.

2- Tokenize words in the transcription column
```{r}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes (x = n, y = fct_reorder(word,n))) + geom_col()

```
The word patient seems to be important, but we observe a lot of stop words.

3- Remove stopwords from visualization
```{r}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  anti_join(stop_words, by = c("word")) %>%

  #exclude anything that is a number b/w 0-9 
  filter(!grepl(pattern = "^[0-9]+$", x= word)) %>% 
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes (x = n, y = fct_reorder(word,n))) + geom_col()
```
4- Tokenize into bigrams and trigrams
```{r}
mtsamples %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes (x = n, y = fct_reorder(bigram,n))) + geom_col()
#do not remove stop words when you dealing with bigrams

```
Using bigrams is not very infomative, let's try trigrams
```{r}
mtsamples %>%
  unnest_ngrams(trigram, transcription, n=3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes (x = n, y = fct_reorder(trigram,n))) + geom_col()
```
There are several more interesting phrases "the operating room", "prepped and draped", tolerated the procedure".

5- Count words before and after a word you've chosen
```{r}
bigrams <- mtsamples %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  filter((word1 == "patient") | (word2 == "patient"))

bigrams %>%
  filter (word1 == "patient") %>%
  select (word1, word2) %>%
  count (word2, sort = TRUE)

bigrams %>%
  filter (word2 == "patient") %>%
  select (word1, word2) %>%
  count (word1, sort = TRUE)

```
Since we are looking at single words agaiin it is a good idea to treat these as single tokens. So let's remove the stopwords and numbers.
```{r}
bigrams %>%
  filter(word1 == "patient") %>%
  filter(!(word2 %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", word2)) %>%
  count(word2, sort=TRUE) %>%
  top_n(10) 
  #knitr::kable()

bigrams %>%
  filter(word2 == "patient") %>%
  filter(!(word1 %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", word1)) %>%
  count(word1, sort=TRUE) %>%
  top_n(10) 
  #knitr::kable()
```

There are 11 words that come before the word "patient" and 10 words that come after.

6- Which words are most used in each specialty? Remove stopwords. What are the 5 most used words?
```{r}
mtsamples %>%
  unnest_tokens(word, input = transcription) %>%
  group_by(medical_specialty) %>%
  count(word, sort = TRUE) %>%
  filter(!(word %in% stop_words$word) & !grepl("^[0-9]+$", word)) %>%
  top_n(5) %>%
  arrange(medical_specialty, n) %>%
  knitr::kable()

```
The 5 most used words per specialty are shown in the above chart.
