---
title: "Lab 4 AT 9-17-21"
output: html_document
- always_allow_html: true 
+ always_allow_html: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

1. Read in the data 
```{r}
library(data.table)
library(tidyverse)
```


```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

2. Prepare the data 

```{r}
met <- met[temp>=-17]
met[,range(temp)]
met[,range(rh, na.rm=TRUE)]
met[,range(wind.sp, na.rm=TRUE)]
met[,range(vis.dist, na.rm=TRUE)]
met[,range(dew.point, na.rm=TRUE)]
met[,range(lat, na.rm=TRUE)]
met[,range(lon, na.rm=TRUE)]
met[,range(elev, na.rm=TRUE)]

met[elev == 9999.9, elev := NA]

met[, ymd := as.Date(paste(year, month, day, sep = ("-")))]

met[, table(week(ymd))]
met <- met[ week(ymd) ==31]

met_avg <- met[, .(
  temp = mean (temp, na.rm = TRUE),
  rh = mean (rh, na.rm = TRUE),
  wind.sp = mean (wind.sp, na.rm = TRUE),
  vis.dist = mean (vis.dist, na.rm = TRUE),
  dew.point = mean (dew.point, na.rm = TRUE),
  lon       = mean (lon, na.rm = TRUE),     
  lat       = mean (lat, na.rm = TRUE),
  elev       = mean (elev, na.rm = TRUE), USAFID
), by = "USAFID"]

```

```{r}
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98, region := "Southeast"]

met_avg[, region2 := fifelse(lat >= 39.71 & lon <= -98, "Northwest",
        fifelse(lat < 39.71  & lon <= -98, "Southwest",
                fifelse(lat >= 39.71 & lon > -98, "Northeast",
                        fifelse(lat < 39.71  & lon > -98, "Southeast", NA_character_))))]

met_avg[, table(region, region2, useNA= "always")]
met_avg[, elev_cat := fifelse(elev >252, "high", "low")]

#delete extra column
met_avg <- met_avg[, -1]
```

3. Use geom_violin to examine the wind speed and dew point temperature by region

```{r}
ggplot(met_avg, mapping = aes(y= wind.sp, x=1)) + 
  geom_violin() + 
  facet_grid(~region)
```
4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region
```{r}

```


6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
```{r}
ggplot(
  met_avg[!is.na(wind.sp) & !is.na(dew.point)], 
  mapping = aes(x = region, y = wind.sp)) + 
    stat_summary(fun.data = "mean_sdl") +
    stat_summary(fun.data = "mean_sdl", geom = "errorbar")
```
7. Make a map showing the spatial trend in relative h in the US

```{r}
library(leaflet)

temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
leaflet(met_avg) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(round(rh,2), ' rh'), color = ~ temp.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=temp.pal, values=met_avg$rh,
          title='Temperature, C', opacity=1)
```

